Title Goes Here

### 2024-03-05 analysis/paper/paper.qmd
topic and geo foci plots

# how many times each location mentioned?
jobdata_geo_year %>%
  group_by(name) %>%
  summarise(n = sum(value)) %>%
  arrange(desc(n)) %>%
  ggplot() +
  aes(reorder(name, n),
      n)+
  geom_col() +
  theme_minimal() +
  coord_flip()

@@

ggplot() +
  geom_line(data =  jobdata_geo_year_tally,
            aes(year_ad_posted,
                 prop,
                 group = name,
                 colour = name) ) +
   geom_point(data = jobdata_geo_year_tally_max,
              aes(year_ad_posted,
                  prop,
                  group = name,
                  colour = name),
              size = 3) +
   geom_text_repel(data = jobdata_geo_year_tally_max,
             aes(year_ad_posted,
                 prop ,
                 group = name,
                 colour = name,
                 label = name),
             nudge_y = 0.01,
             bg.color = "white",
             bg.r = 0.25,
             size = 3) +
  xlab("Year") +
  ylab("Proportion of all ads") +
  theme_minimal()

@@

>>>>

# how many times each location mentioned?
jobdata_geo_year %>%
  group_by(name) %>%
  summarise(n = sum(value)) %>%
  arrange(desc(n)) %>%
  ggplot() +
  aes(reorder(name, n),
      n)+
  geom_col() +
  xlab("") +
  theme_minimal() +
  coord_flip()

@@

ggplot() +
  geom_line(data =  jobdata_geo_year_tally,
            aes(year_ad_posted,
                 prop,
                 group = name,
                 colour = name),
            size = 2,
            position=position_jitter(w=0.05, h=0.0)) +
   geom_point(data = jobdata_geo_year_tally_max,
              aes(year_ad_posted,
                  prop,
                  group = name,
                  colour = name),
              size = 5) +
   geom_text_repel(data = jobdata_geo_year_tally_max,
             aes(year_ad_posted,
                 prop ,
                 group = name,
                 colour = name,
                 label = name),
             nudge_y = 0.01,
             bg.color = "white",
             bg.r = 0.25,
             size = 3) +
  xlab("Year") +
  ylab("Proportion of all ads") +
  theme_minimal()

@@

```{r}

# topical focus by year

topical_foci <-
  read_sheet("https://docs.google.com/spreadsheets/d/1AHq49pIyChcgJ7rawe6KMWkdIBXydCamvg8Jslob8Ec/edit#gid=0",
             sheet = "topic")

topical_foci_clean <-
  map(
    str_split(topical_foci$`From the data`, ";"),
    ~.x %>%
      str_squish() %>%
      stri_remove_empty() %>%
      str_to_lower)

jobdata_topic <-
  jobdata %>%
  select(topical_focus_of_position) %>%
  mutate(topical_focus_of_position = str_to_lower(topical_focus_of_position))

jobdata_topic <-
  # add one column for each topic in our categories
  cbind(jobdata_topic,
        setNames( lapply(topical_foci$Category, function(x) x=NA),
                  topical_foci$Category) )

for(i in 1:length(topical_foci$Category)){

  this_topic <- topical_foci$Category[i]

  # create the pattern to search for
  x <- paste0(topical_foci_clean[[i]], collapse = "|")

  # do the search through all the job ads for that pattern
  y <- str_detect(jobdata_topic$topical_focus_of_position,
                  x)

  # assign back to our data frame in the appropriate location column
  jobdata_topic[, this_topic] <- y

jobdata_topic_year <-
  jobdata %>%
  bind_cols(jobdata_topic) %>%
  select(year_ad_posted,
         topical_foci$Category) %>%
  pivot_longer(-year_ad_posted) %>%
  drop_na()

# how many times each location mentioned?
jobdata_topic_year %>%
  group_by(name) %>%
  summarise(n = sum(value)) %>%
  arrange(desc(n)) %>%
  ggplot() +
  aes(reorder(name, n),
      n)+
  geom_col() +
  xlab("") +
  theme_minimal() +
  coord_flip()

# explore trends over time. put a point on the max year
jobdata_topic_year_tally <-
  jobdata_topic_year %>%
  # exclude those with <20 ads
  filter(!name %in% c("Digital Archaeology",
                      "Pleistocene archaeology",
                      "Mesoamerican Archaeology",
                      "Biological anthropology",
                      "Archaeological theory",
                      "Evolutionary anthropology",
                      "North American archaeology"
  )) %>%
  group_by(year_ad_posted,
           name) %>%
  summarise(n = sum(value)) %>%
  mutate(prop = n / sum(n))

jobdata_topic_year_tally_max <-
  jobdata_topic_year_tally %>%
  group_by(
    name ) %>%
  filter(prop == max(prop))

ggplot() +
  geom_line(data =  jobdata_topic_year_tally,
            aes(year_ad_posted,
                prop,
                group = name,
                colour = name),
            size = 2,
            position=position_jitter(w=0.05, h=0.0)) +
  geom_point(data = jobdata_topic_year_tally_max,
             aes(year_ad_posted,
                 prop,
                 group = name,
                 colour = name),
             size = 10) +
  geom_text_repel(data = jobdata_topic_year_tally_max,
                  aes(year_ad_posted,
                      prop ,
                      group = name,
                      colour = name,
                      label = name),
                  nudge_y = 0.01,
                  bg.color = "white",
                  bg.r = 0.25,
                  size = 3) +
  xlab("Year") +
  ylab("Proportion of all ads") +
  theme_minimal()

ggsave(here("analysis",
            "figures", 
            "fig-topic-focus-by-year.png"),
       bg ="white")
